### 2020-01-28 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
мелкие правки: nonliner -> non-convex часть текста сделан курсивом

Among the approaches which have been developed for solving MCO problems, lexicographic optimization methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}. Iterative methods represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions. An actively appropriated approach consists in developing and applying evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}. Among the widely used endeavors in solving MCO problems is the scalarization approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g. \cite{c2,c14}).

This work is devoted to solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems. In such applications, efficiency criteria may have a complex multiextremal form, and calculating criteria values may require large amounts of computations. In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of great computational complexity.

@@

This paper proposes a solution to the time-consuming class of MCOlex problems by using an approach based on the following main points. First of all, the problem is reduced to solving a sequence of global optimization problems with nonlinear constraints (GCO) \cite{c2,c14}. Then, efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}. And finally, the search information obtained in the process of solving the MCOlex problem is fully utilized. In general, the developed approach allows us to significantly reduce the amount of executed computations, up to just a few iterations when searching for new efficient decisions.

The further structure of the paper is as follows. Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with nonlinear constraints. Section \ref{sec:3} presents the basics of the approach that has been developed: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations. Section \ref{sec:4} contains the results of numerical experiments. The paper concludes by discussing the results that were obtained and suggests some main directions for further research.

@@

This approach is also widely known as the method of successive concessions (MSC) \cite{c3,c4,c5}. The choice of feasible concessions $\delta_i$, $1 \leq i < s$, in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved. At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with nonlinear constraints.

@@

The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with nonlinear constraints. Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality. The proposed approach for solving such problems is based on three key ideas:
\begin{itemize}
  \item reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
  \item using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
  \item repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
\end{itemize}

@@

The general computational scheme of the global search algorithm for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.

>>>>

Among the approaches which have been developed for solving MCO problems, \textit{lexicographic optimization} methods can be singled out, when some ordering of criteria by importance is implemented and optimization is carried out based on their ordering \cite{c3}. \textit{Iterative methods} represent another approach \cite{c6,c10}, when the decision maker is actively involved in the process of selecting decisions. An actively appropriated approach consists in developing and applying evolutionary algorithms based on the simulation of certain natural phenomena to solve MCO problems \cite{c10,c11,c12,c13}. Among the widely used endeavors in solving MCO problems is the \textit{scalarization} approach, which applies methods involving the convolution of criteria to a single scalar criterion (see, e.g. \cite{c2,c14}).

This work is devoted to solving lexicographic multicriteria optimization (MCOlex) problems that arise when designing complex technical objects and systems. In such applications, efficiency criteria may have a complex \textit{multiextremal} form, and calculating criteria values may require \textit{large amounts of computations}. In such conditions, finding even one efficient decision requires significant calculations, while determining several (or a whole set) of efficient decisions becomes an issue of \textit{great computational complexity}.

@@

This paper proposes a solution to the time-consuming class of MCOlex problems by using an approach based on the following main points. First of all, the problem is reduced to solving a sequence of global optimization problems with non-convex constraints (GCO) \cite{c2,c14}. Then, efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization are applied for solving the GCO problems \cite{c17,c18}. And finally, the search information obtained in the process of solving the MCOlex problem is fully utilized. In general, the developed approach allows us to significantly reduce the amount of executed computations, up to just a few iterations when searching for new efficient decisions.

The further structure of the paper is as follows. Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems -- whose solution is reduced to solving a sequence of global optimization problems with non-convex constraints. Section \ref{sec:3} presents the basics of the approach that has been developed: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations. Section \ref{sec:4} contains the results of numerical experiments. The paper concludes by discussing the results that were obtained and suggests some main directions for further research.

@@

This approach is also widely known as the method of successive concessions (MSC) \cite{c3,c4,c5}. The choice of feasible concessions $\delta_i$, $1 \leq i < s$, in the computational scheme (\ref{eq:7}) allows one to obtain any efficient decision $\widetilde{y}^* \in P_{lex}(f,D)$ and take into account the peculiarities of the MCOlex problem being solved. At the same time, this approach makes it necessary, at each stage of scheme (\ref{eq:7}), to solve more complex global optimization problems with non-convex constraints.

@@

The extended formulation of lexicographic multicriteria optimization problems (\ref{eq:7})--(\ref{eq:8}) involves multiple solutions of multiextremal optimization problems with non-convex constraints. Problems of this type are computationally time-consuming and subject to the ``curse of dimensionality'': computational complexity increases exponentially with increasing dimensionality. The proposed approach for solving such problems is based on three key ideas:
\begin{itemize}
  \item reduction of multidimensional problems of multiextremal optimization to problems of one-dimensional global search by using mappings based on Peano curves;
  \item using the index method of global optimization, which allows solving optimization problems with constraints without using penalty functions;
  \item repeated use of search information obtained in the process of computing, for solving multiple global optimization problems.
\end{itemize}

@@

The general computational scheme of the \textit{global search algorithm} for multiextremal optimization problems with non-convex constraints \footnote{This method is also known as the index method -- see \cite{c18}.} (AGCS) is as follows \cite{c18}.
